---
title: "Lab-07-Replication"
author: "Branson Fox"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook replicates the results of lab 07.

## Dependencies
This notebook requires the following packages to load, clean and export our data.

```{r load-packages}
# tidyverse packages
library(readr) # import tabular data
library(dplyr) # data manipulation

# other packages
library(here) # file path management
library(sf)   # methods for spatial data
library(measurements) # unit conversion
```

## Load Data
This notebook requires three pieces of raw data from the lecture 08 repository.

```{r load-data}
# spatial data
counties <- st_read(here("data", "MO_BOUNDARY_Counties", "MO_BOUNDARY_Counties.shp"), stringsAsFactors = FALSE)
countyPop <- st_read(here("data", "MO_DEMOS_CountyPop", "MO_DEMOS_CountyPop.shp"), stringsAsFactors = FALSE)

# tabular data
countyDisability <- read_csv(here("data", "MO_DEMOS_CountyDisability/MO_DEMOS_CountyDisability.csv"))
```

## Part 1: Data Wrangling
### Question 1
We'll begin by creating a new variable that adds together rates of disability by poverty to find the total rate of disability between 0 and 150% of the poverty line.
```{r p1-q1}
countyDisability %>%
  mutate(
    TotalDisability =  under50_dis + btwn50_99_dis + btwn100_150_dis
  ) %>%
  select(GEOID, TotalDisability) -> countyDisability
```

### Question 2
Now, we'll remove the `NAME` column from the county population data and convert it to a data.frame.
```{r p1-q2}
countyPop <- select(countyPop, -NAME)
```

First, observe that it is of class sf:
```{r q2 sf}
class(countyPop)
```

Now, to remove the geometry:
```{r q2 rm geo}
st_geometry(countyPop) <- NULL
```

Now, see that it is no longer of class `sf`
```{r q2 sf2}
class(countyPop)
```

### Question 3
Next, we'll join the county population and disability data. Because of the way a shapefile and csv are read in differently, we need to change the character format of countyPop GEOID to be numeric.
```{r p1-q3}
# change variable class
countyPop <- mutate(countyPop, GEOID = as.numeric(GEOID))
# then to join as usual
countyFull <- left_join(countyDisability, countyPop, by = "GEOID")
```

### Question 4
Next, we'll calculate the proportion of the county that has a dissability and is between 0 and 150% of the poverty line out of the total county population.
```{r p1-q4}
countyFull <- mutate(countyFull,
                     disProp = TotalDisability/totalPop)
```

### Question 5
Our data is in the correct order now, but if you need to rearrange the order of variables, use the `select()` function and manually rearrange them.
```{r p1-q5}
countyFull <- select(countyFull, GEOID, TotalDisability, totalPop, disProp)
```

### Question 6
Now, we want to remove all of the columns from the county data except for `GEOID` variable we can join by. Remember that sf geometries are "sticky" or in other words they will remain unless explicitly dropped.
```{r p1-q6}
counties <- select(counties, GEOID)
```

### Question 7
Now, we will join our `countyFull` data with `counties`. Once again, we will change the GEOID from character to numeric in order to match.
```{r p1-q7}
# change class of GEOID
counties <- mutate(counties, GEOID = as.numeric(GEOID))
# join as usual
countyFull <- left_join(counties, countyFull, by = "GEOID")
```

### Question 8
The last modification we'll make to the data is calculating the square kilometer area per county. First we use `st_area()` to get the area in square meters. We have to force to numeric using `as.numeric()` or else there will be trouble converting later. Then, we use the `conv_unit()` function from `measurements` to convert this to square km.
```{r p1-q8}
# to get area from geometry
countyFull <- mutate(countyFull, area = as.numeric(st_area(geometry)))
# then to convert to square km
countyFull <- mutate(countyFull, area = conv_unit(area, "m2", "km2"))
```

### Question 9
Finally, we will write the cleaned data as a shapefile to the `data/` folder. In this example, we created a new folder called `MO_CountyDisabilityTotal` as a subfolder of `data/`.
```{r p1-q9}
# an example of creating a folder by R code
dir.create(here("data", "MO_CountyDisabilityTotal"))
# then regular saving of our data
st_write(countyFull, here("data", "MO_CountyDisabilityTotal", "MO_CountyDisabilityTotal.shp"))
```

## Part 2: Geodatabase Creation

This part of the lab is completed in ArcGIS Pro.
